### 2017-07-17 sheets/talk-11/sheet-11.tex
Minor reformulations
Add first solutions to sheet 11

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

@@

\begin{exercise}
  Define the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
  the presheaf model.
\end{exercise}
\begin{answer}
  $\mathit{Ty}_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set
  with some objects (possibly empty).
  \[
    \mathit{Ty}_F(\Gamma) := \{A_\Gamma, B_\Gamma, \dots\}
  \]
  $\mathit{Ter}_F(\Gamma, A)$ is then an element of the second projection of
  $F(\Gamma)$, that is a member of the family of sets indexed over
  $\mathit{Ty}_F(\Gamma)$, which has A as its index and also may be empty.
  \[
    \mathit{Ter}_F(\Gamma, A) := \{a_A, b_A, \dots\}_A
  \]
\end{answer}

\begin{exercise}
  Let $(\cat C, \mathcal F)$ be a category with families.  Check that the
  following square is a pullback in $\cat{C}$:
  \[
    \begin{tikzcd}
      \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
      \Delta \arrow{r}{\sigma} & \Gamma
    \end{tikzcd}
  \]
  where $\Delta.A \vdash q : A p_\Delta$
\end{exercise}

@@

  Then verify that this construction imposes a functor $\int_\cat{C} \from
  \mathit{Psh}(\cat{C}) \to \CAT$.
\end{exercise}
\begin{answer}
\end{answer}

@@

  \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.

\end{exercise}

@@

  \emph{Hint}: The definition is \emph{almost} identical to the one given
  above. But what do you need to change? To recall the presheaf model, read
  section $1.2$ in Huber's thesis.
\end{exercise}
\begin{answer}
  Section $1.2.2$.
\end{answer}

@@

  Define a length function of type $\mathit{List}(A) \to \mathbb{N}$ and define
  a type $\mathit{Vec}_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
  using lists, the identity type and the $\Sigma$-type.
\end{exercise}

@@

  \emph{Hint}: define using $R^\mathbb{N}$ a function $f \from \mathbb{N} \to
  U$ such that $\diamond \vdash f 0 = \hat{1}: U$ and $\diamond \vdash
  f(Suc(0)) = \hat{0} : U$
\end{exercise}

>>>>

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
\usepackage{semantic}

@@

\begin{exercise}
  Give the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
  the presheaf model.
\end{exercise}
\begin{answer}
  \[
    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\Gamma)}
  \]
  \[
    \mathit{Ter}_F(\Gamma, A) := \{(a\rho \in A\rho)_{I \in \cat{C}, \rho \in \Gamma(I)} | \forall f \from J \to I, a\rho f = a(\rho f) \}
  \]
\end{answer}

\begin{exercise}
  Let $(\cat C, \mathcal F)$ be a category with families.  Check that the
  following square is a pullback in $\cat{C}$:
  \[
    \begin{tikzcd}
      \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
      \Delta \arrow{r}{\sigma} & \Gamma
    \end{tikzcd}
  \]
  where $\Delta.A \vdash q : A p_\Delta$
\end{exercise}
\begin{answer}
  Let $\Delta'$ be a context and $p' \from \Delta' \to \Delta$ and $\sigma'
  \from \Delta' \to \Gamma.A$ be two morphisms such that $\sigma \circ p' =
  p_\Gamma \circ \sigma'$. Therefore the diagram
  \[
    \begin{tikzcd}
      \Delta' \arrow[bend right=30]{ddr}{p'} \arrow[dashed]{dr}{\delta} \arrow[bend left=30]{drr}{\sigma'}&   &\\
          & \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
          & \Delta \arrow{r}{\sigma} & \Gamma
    \end{tikzcd}
  \]
  (without $\delta$) is commutative. We must show that there exists a unique
  $\delta \from \Delta' \to \Delta.A\sigma$ such that $(\sigma \circ p_\Delta,
  q) \circ \delta = \sigma'$ and $p_\Delta \circ \delta = p'$.

  We can decompose $\sigma'$ as $\sigma' = (p_\Gamma, q_\Gamma) \circ \sigma' =
  (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma \circ p', M)$, where $M
  \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash A(\sigma \circ p')
  ~\text{type}$). Let $\delta$ be morphism
  \[
    \delta \equiv (p', M) \from \Delta' \to \Delta.A\sigma
  \]
  which is suitable since $p_\Delta \circ \delta = p'$ and
  \[
    (\sigma \circ p_\Delta, q) \circ \delta =
    (\sigma \circ p_\Delta \circ \delta, q\delta) =
    (\sigma \circ p', q_\Gamma\sigma') =
    (p_\Gamma \circ \sigma', q_\Gamma\sigma') =
    (p_\Gamma, q_\Gamma) \circ \sigma' =
    \sigma'
  \]

  For uniqueness assume morphism $\delta' \from \Delta' \to \Delta.A\sigma$
  such that $(\sigma \circ p_\Delta, q) \circ \delta' = \sigma'$ and $p_\Delta
  \circ \sigma' = p'$. As before we have $\delta' = (p_\Delta \circ \delta',
  q\delta') = (p', N)$ where $N \equiv q\delta'$. The following equalities
  hold:
  \[
    N = q\delta' = q_\Gamma(\sigma \circ p_\Delta \circ \delta', q\delta') = q_\Gamma((\sigma \circ p_\Delta, q) \circ \delta') = q_\Gamma\delta' = M
  \]
  and therefore $\delta' = \delta$.
\end{answer}

@@

  Then verify that this construction imposes a functor $\int_\cat{C} \from
  \mathit{Psh}(\cat{C}) \to \CAT$.
\end{exercise}
\begin{answer}
  Indeed, the operation of $\sigma$ on such objects is a precomposition
  satisfying the naturality axiom:
  \[
    \int_\cat{C}\sigma(I, \rho) := (I, \sigma_I \rho)
  \]

  Let's check that $\int_\cat{C}$ indeed takes us into a category.
  \begin{itemize}
    \item objects are pairs $(I, \rho)$ where $I \in \cat{C}$ and $\rho \in \Gamma(I)$
    \item a morphism $(J, \rho') \to (I, \rho)$ is a morphism $f \from J \to I$
      in $\cat{C}$ such that $\rho' = \rho f$.
  \end{itemize}
  We need to check that we have composition and identities.

  With our definition of objects, we can consider first and second projections
  of the pairs separately. We can observe, that elements of first projection
  are elements in $\cat{C}$ which have required identities and composition
  under morphisms in $\cat{C}$ (which are also our morphisms). Then what
  remains is the second projection. Identities and composition are exactly our
  conditions on presheaves which also hold.

  Now we know the operation of $\int_\cat{C}$ on objects and arrows in
  $Psh(\cat{C})$.
\end{answer}

@@

  \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.
\end{exercise}
\begin{answer}
  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we the elements of $\mathit{Ty}(\Gamma)$:
  \[
    (A \to A)\rho = \{w | w_f \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
  \]
  And required equalities hold trivially for any $g \from K \to J$: $((\lambda
  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.

  Then we give a term by the following equation:
  \[
    ((\lambda q) \rho)_f u = q(\rho f, u) = u
  \]
\end{answer}

@@

  \emph{Hint}: The definition is \emph{almost} identical to the one given
  above. But what do you need to change? To recall the presheaf model, read
  section $1.2$ in Huber's thesis.
\end{exercise}
\begin{answer}
  \[
    (\Sigma AB)\rho := \{(a,b) | a \in A\rho \land b \in B(\rho, a)\}
  \]
\end{answer}

@@

  Define a length function of type $\mathit{List}(A) \to \mathbb{N}$ and define
  a type $\mathit{Vec}_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
  using lists, the identity type and the $\Sigma$-type.
\end{exercise}
\begin{answer}
  First let's define the type formers:
  \[
    \inference[List-F]{\Gamma \vdash A ~\text{type}}{\Gamma \vdash \mathbf{List_A} ~\text{type}}\quad
    \inference[List-I-nil]{\Gamma \vdash A ~\text{type}}{\Gamma \vdash \mathit{nil_A} : \mathbf{List_A}}\quad
    \inference[List-I-cons]{\Gamma \vdash a : A \quad \Gamma \vdash x : \mathbf{List_A}}{\Gamma \vdash \mathit{Cons(a, x)} : \mathbf{List_A}}
  \]

  \[
    \inference[List-E]{
      \Gamma, l: \mathbf{List_A} \vdash B ~\text{type}\\
      \Gamma \vdash H_{nil} : B[nil_A/l]\\
      \Gamma, a : A, l : \mathbf{List_A}, x: B \vdash H_{cons} : B[Cons(a, l)/l]\\
      \Gamma \vdash l : \mathbf{List_A}
    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, l) : B[l/l]}
  \]

  \[
    \inference[List-C-nil]{
      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, nil) : B[nil/l]
    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B]H_{cons}, nil) = H_{nil} : B[nil/l]}
  \]

  \[
    \inference[List-C-cons]{
      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(a, l)) : B[Cons(a, l)/l]
    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(a, l)) = D}
  \]
  where $D = H_{cons}[a/a, l/l, R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, l)/x] : B[Cons(a, l)/l]$.

  Now we can define the length function:
  \[
    \mathit{Length(l) \vcentcolon= R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]\mathbb{N}}(0, [a : A, l: \mathbf{List_A}, x : \mathbb{N}] Suc(x), l) : \mathbb{N}}
  \]

  And finally the $\mathit{Vec_A(n)}$ type:
  \[
    \mathit{Vec_A(n)} \vcentcolon= \Sigma l: \mathbf{List_A}. Id_\mathbb{N}(n, \mathit{Length}(l))
  \]
\end{answer}

@@

  \emph{Hint}: define using $R^\mathbb{N}$ a function $f \from \mathbb{N} \to
  U$ such that $\diamond \vdash f 0 = \hat{1}: U$ and $\diamond \vdash
  f(Suc(0)) = \hat{0} : U$
\end{exercise}
\begin{answer}
  \[
    \inference[U-0]{\vdash \Gamma~\text{ctxt}}{\Gamma \vdash \hat{0} : U}\quad
    \inference[U-0-Ty]{\Gamma \vdash \hat{0}: U}{\Gamma \vdash El(\hat{0}) = 0~\text{type}}
  \]

  \[
    \inference[U-1]{\vdash \Gamma~\text{ctxt}}{\Gamma \vdash \hat{1} : U}\quad
    \inference[U-1-T]{\Gamma \vdash \hat{1} : U}{\Gamma \vdash El(\hat{1}) = 1~\text{type}}
  \]

  \[
    f(n) = R^\mathbb{N}_{[n:\mathbb{N}]U}(\hat{1}, [n: \mathbb{N}, x: U]\hat{0}, n)
  \]

  Now to get an inhabitant of that type let's use $\mathit{Subst}$:
  \[
    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, El(f(0))) : El(f(Suc(0)))
  \]
  Since the inhabitant to $El(f(0)) = El(\hat{1}) = 1~\text{type}$, namely
  $\star$, is always present by the formation rules, and $El(f(Suc(0))) =
  0~\text{type}$ we obtain an inhabitant to the required type.
\end{answer}
